---
title: "Insurance Quotes"
author: "Anthony Arroyo, Seung min song, Alice A. Friedman,Ivan Tikhonov, Charles Ugiagbe"
date: "2023-05-05"
output:
  pdf_document: default
  html_document: default
---

  
```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(here)
library(reshape2)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ModelMetrics)
library(readr)
library(e1071)
library(purrr)
library(tidyr)
library(corrplot)
library(FactoMineR)
library(VIFCP)
library(knitr)
library(kableExtra)
library(Hmisc)
library(pROC)
library(binr)


rawData = read.csv(file="https://raw.githubusercontent.com/IvanGrozny88/DATA-621-HW-4/main/insurance_training_data.csv")
rawTestData = read.csv(file="https://raw.githubusercontent.com/IvanGrozny88/DATA-621-HW-4/main/insurance-evaluation-data.csv")

```


# Abstract

We discuss the challenges from evaluating insurance data, asserting whether a car is involved in a crash and then how expensive the claim would be. Mainly, we discuss imputation but our end models are not able to effectively predict car crashes due to overfitting and high false positives. Our linear regression models pull some insights regarding the effects of marital status and others on claim costs.

# Data Exploration & Preparation

The data has 23 features and 2 target variables that we wish to predict. As part of our predictive analytics, first we will decide if the car that belongs to a license plate has been in a crash before. Second, we will determine how much the cost of the crash was if they were.

- TARGET_FLAG: Was Car in a crash? 1=YES 0=NO None
- TARGET_AMT: If car was in a crash, what was the cost

There are some general myths related to driving that may skew our view about what we expect. How much of it is true? Personally, I think that senior drivers get into more crashes than those younger but what if that's not the case? Provided below is a quick data dictionary of the features we will be measuring.

- AGE: Age of Driver
- BLUEBOOK: Value of Vehicle
- CAR_AGE: Vehicle Age
- CAR_TYPE: Type of Car
- CAR_USE: Vehicle Use
- CLM_FREQ: # Claims (Past 5 Years)
- EDUCATION: Max Education Level
- HOMEKIDS: # Children at Home
- HOME_VAL: Home Value
- INCOME: Income
- JOB: Job Category
- KIDSDRIV: # Driving Children
- MSTATUS: Marital Status
- MVR_PTS: Motor Vehicle Record Points
- OLDCLAIM: Total Claims (Past 5 Years)
- PARENT1: Single Parent
- RED_CAR: A Red Car
- REVOKED: License Revoked (Past 7 Years)
- SEX: Gender
- TIF: Time in Force
- TRAVTIME: Distance to Work
- URBANICITY: Home/Work Area
- YOJ: Years on Job

The top of the dataset is shown below, where we can see a smattering of isues such as untidy data and even a missing value in one of the income observations.

```{r}
#this step is necessary in order to analyze data as it is not clean
currencyconv = function(input) {
  out = sub("\\$", "", input)
  out = as.numeric(sub(",", "", out))
  return(out)
}

# Replace spaces with underscores
underscore = function(input) {
  out = sub(" ", "_", input)
  return(out)
}


rawData = as.tbl(rawData) %>% 
  mutate_at(c("INCOME","HOME_VAL","BLUEBOOK","OLDCLAIM"),
            currencyconv) %>% 
  mutate_at(c("EDUCATION","JOB","CAR_TYPE","URBANICITY"),
            underscore) %>% 
  mutate_at(c("EDUCATION","JOB","CAR_TYPE","URBANICITY"),
            as.factor) %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG))

#check data
summary(rawData) %>% kable() %>% kable_styling()
```

## Tidy Data

After carefully examining the data in the raw data frame, it was found that certain variables contained unnecessary characters such as dollar signs and garbage data. To address this issue, regular expression (regex) was used to remove these unwanted characters from the variables. By doing so, the data is now cleaner and more suitable for analysis.

Additionally, there are several nominal categories that are yes and no questions. We reduced this to binary where 1=yes and2=no using regex again to further simplify the process.

Finally, we convert variables to type factor or numeric, as appropriate.

```{r clean-funcs}
### Akin to dplyr::mutate_at but does not throw error ("silent") on missing cols and applies a function
SilentMutateAt <- function(df, dirtyCols = c("Sample", "Text"), FUN, ...) {
  ind = colnames(df) %in% dirtyCols
  df[, ind] = sapply(df[, ind], FUN = FUN, ... = ...)
  
  return(df)
}

### Rearranges parameters of gsub to be more friendly to pipe operator
GsubPipe <- function(x, pattern, replacement, ...) {
  gsub(pattern, replacement, x, ...)
}

### Replaces "yes/no" questions with an integer where yes=1 and no=0
CleanYesNo <- function(df, trace = FALSE, threshold = 0.5) {
  yesNoCols = sapply(df, function(x) { 
    unique(x) %>% 
      grepl("yes|no|^y$|^n$", ., ignore.case = T) %>% 
        { sum(.) / length(.) } %>%
          { . >= threshold }
  }) %>%
    .[.] %>%
      names(.)
  
  result = SilentMutateAt(df, yesNoCols, function(x) {
    truthy = as.integer(grepl("yes|^y$", x, ignore.case = T))
    truthy = truthy - as.integer(grepl("no|^n$", x, ignore.case = T))
    
    truthy[truthy == 0L] = NA
    truthy[truthy == -1L] = 0L
    
    truthy
  })
  
  if (trace) {
    return(list(result, yesNoCols))
  }
  
  return(result)
}

### Pre-processing data pipeline for insurance data; cleans up strings
CleanInsuranceStrings <- function(df) {
  SilentMutateAt(df, colnames(df)[sapply(df, is.character)], GsubPipe, "\\$|z_|<|,", "") %>% 
    CleanYesNo(.) %>%
      # could also use readr::parse_number here
      SilentMutateAt(c("INCOME", "HOME_VAL", "BLUEBOOK", "OLDCLAIM"), as.numeric) %>% 
        ### Coerces blank strings to NA
        SilentMutateAt(., "JOB", function(x) { x[x == ""] = "Unspecified" ; x }) %>%
          ### Converts strings to factors
          dplyr::mutate_at(., colnames(.)[sapply(., is.character)], as.factor)
}
```

```{r init}
#runs functions
cleanData <- CleanInsuranceStrings(rawData)
cleanTestData <- CleanInsuranceStrings(rawTestData)
imputeData = CleanInsuranceStrings(rawData)

#reference: https://statisticsglobe.com/convert-character-to-factor-in-r

summary(cleanData)
head(cleanData)
```

## Visulization of the data set

Our features span a multitude of different distributions. Many of them are normal, others are binomial, and others are nominal. How we handle each feature will thus influence future analysis. 

```{r}

### Plots the distribution of each variable
# See common-helpers.R
ntrain<-select_if(rawData, is.numeric)
ntrain %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density() 
```


## Visualization of Correlation

According to this correlation matrix, all of the features are not strongly correlated to each other. With this in mind, we will go forward and assume that potentially every feature is a principal component and we will not exclude any until further downstream analysis is conducted.

```{r corr}
### Develops a correlation plot for all numeric values
# See common-helpers.R
rawDatanum <- dplyr::select_if(rawData, is.numeric)

rcorr(as.matrix(rawDatanum))
corrplot(cor(rawDatanum), method="square")
```

## Missing values
The raw data frame contains `r sum(colSums(is.na(rawData)))`` missing values, representing `r sum(!complete.cases(rawData))` out of `r nrow(rawData)` observations.

```{r missing, eval=FALSE}
### Counts values with missing data
#I don't understand this line, but this does not agree with the next chunk
# AA: i got the same result on my side but the other version is more clear
#nrow(rawData[is.na(rawData),]) 

### total missing values
colSums(is.na(rawData)) %>% sum()

### Counts rows with missing data
sum(!complete.cases(rawData)) #observations with missing data
```

The following variables in the dataset have missing values: **TARGET_FLAG**, **TARGET_AMT**, **AGE**, **YOJ**, and **CAR_AGE**. 

```{r missing-cols}
#this function adds up the rows with missing values to get a count of NAs by feature
colSums(is.na(rawData)) 
```
A good rule of thumb is to drop any feature with more than 80% missing values -- that is not the case with any features here. The remaining options are to:
  1) Drop rows with missing values
2) Infer missing values

In order to decide which course to take we should first determine if there is any correlation between features with missing values, or if values are missing at random. Because several of these features are numeric, we will first need to convert the features to the appropriate data type in order to determine if there is a correlation.

```{r warning=FALSE}
# TO COME -- Section exploring correlation
# impute data for missing values
# use column mean for calculation

rawData$AGE[is.na(rawData$AGE)] <- mean(rawData$AGE, na.rm=TRUE)
rawData$YOJ[is.na(rawData$YOJ)] <- mean(rawData$YOJ, na.rm=TRUE)
rawData$HOME_VAL[is.na(rawData$HOME_VAL)] <- mean(rawData$HOME_VAL, na.rm=TRUE)
rawData$CAR_AGE[is.na(rawData$CAR_AGE)] <- mean(rawData$CAR_AGE, na.rm=TRUE)

rawData$INCOME[is.na(rawData$INCOME)] <- mean(rawData$INCOME, na.rm=TRUE)

#get complete cases
rawData <- rawData[complete.cases(rawData),]

rawData2<-rawData
 
# # transform data using log for skewed HOMEKIDS, MVR_PTS, OLDCLAIM, TIF, KIDSDRIVE and CLM_FREQ 

rawData$HOMEKIDS <- log(rawData$HOMEKIDS+1)
rawData$MVR_PTS <- log(rawData$MVR_PTS+1)
rawData$TIF <- log(rawData$TIF+1)
rawData$KIDSDRIV <- log(rawData$KIDSDRIV+1)
rawData$CLM_FREQ <- log(rawData$CLM_FREQ+1)


#remove rad per correlation in prior section

rawData <- rawData[, !(colnames(rawData) %in% c("INDEX"))]
# 
# #create variable
rawDatanum <- dplyr::select_if(rawData, is.numeric)

rcorr(as.matrix(rawDatanum))
```

It is important to address these missing values before proceeding with any analysis.


### AGE & TARGET_FLAG

Based on the available data, it seems that YOJ tends to increase with age. There are 6 missing values in the AGE column, which is a trivial enough number that the simplest approach is to drop those observations. Observations that are missing the target cannot be used for training so those will also be dropped.


### YOJ

Build the linear regression model with AGE as the predictor and YOJ as the response and predict the missing values of YOJ based on the values of AGE. Insert the predicted values of YOJ into the original dataframe.

```{r impute-yoj}
# Subset the data with missing values of YOJ
missing_YOJ <- subset(imputeData, is.na(YOJ))

lm_model <- lm(YOJ ~ AGE, data = imputeData)

predicted_YOJ <- predict(lm_model, newdata = missing_YOJ)

rounded_YOJ <- round(predicted_YOJ, digits = 0)
imputeData$YOJ[is.na(imputeData$YOJ)] <- rounded_YOJ


```

### Income

Build the linear regression model with AGE and YOJ as the predictors and INCOME as the response and predict the missing values of INCOME based on the values of AGE and YOJ. insert the predicted values of INCOME into the original dataframe.

```{r impute-income}
# Subset the data with missing values of INCOME
missing_INCOME <- subset(imputeData, is.na(INCOME))

lm_model <- lm(INCOME ~ AGE + YOJ, data = imputeData)

predicted_INCOME <- predict(lm_model, newdata = missing_INCOME)

rounded_INCOME <- round(predicted_INCOME, digits = 0)
imputeData[is.na(imputeData$INCOME),]$INCOME <- rounded_INCOME

```

### Home-val

Build the linear regression model with AGE, YOJ, and INCOME as predictors and HOME_VAL as the response and predict the missing values of HOME_VAL based on the values of AGE, YOJ, and INCOME. Insert the predicted values of HOME_VAL into the original dataframe

```{r impute-home}
# Subset the data with missing values of HOME_VAL
missing_HOME_VAL <- subset(imputeData, is.na(HOME_VAL))


lm_model <- lm(HOME_VAL ~ AGE + YOJ + INCOME, data = imputeData)


predicted_HOME_VAL <- predict(lm_model, newdata = missing_HOME_VAL)

#rounded_HOME_VAL <- round(predicted_HOME_VAL, digits = 0)
imputeData[is.na(imputeData$HOME_VAL),]$HOME_VAL <- predicted_HOME_VAL
imputeData$HOME_VAL <- round(imputeData$HOME_VAL, digits = 0)

head(imputeData$HOME_VAL)
```

### Car Age

To handle the missing value in CAR_AGE, we will use chained imputation since it is difficult to assess. This will also fill any other data that is still missing so we now have produced our final cleaning steps. We may next consider doing manipulations such as normalization, zero centering, and other kinds of feature engineering.

```{r impute-finish}
imputeData = imputeData %>%
  mice::mice(., method = "pmm", printFlag = FALSE) %>%
  mice::complete(.)
trainClean = imputeData
```

## Summary

```{r impute-summary}
summary(dplyr::select(imputeData, -INDEX))
```

## Box-plot

We have 971 male and 1170 female in this study. Male average income is higher than man. However, man has highest car purchase amount. The boxplots show that the medians and interquartile range of the two distributions are very close.

```{r sex-income}
SEX_df <- imputeData %>%
  dplyr::filter(!is.na(INCOME)) %>%
  dplyr::group_by(SEX) %>%
  dplyr::summarise(INCOME = mean(INCOME))

SEX_df
```

```{r plot-sex-income}
ggplot2::ggplot(imputeData) +
  ggplot2::aes(x = SEX, y = INCOME, color = SEX, fill = SEX) +
  ggplot2::geom_bar(data = SEX_df, stat = "identity", alpha = .3) +
  ggrepel::geom_text_repel(ggplot2::aes(label = HOME_VAL), color = "black", size = 2.5, segment.color = "grey") +
  ggplot2::geom_point() +
  ggplot2::guides(color = "none", fill = "none") +
  ggplot2::theme_bw() +
  ggplot2::labs(title = "Income by Sex", x = "Sex", y = "Income")
```




# Model Building

We will each create a model to explore and compare the differences between them. Additionally, we will discuss the key insights that we gain from the modeling process because it will help us draw conclusions about the dataset and potentially point us towards a preferred model.

## Seung's Model logit

Stepwise regression analysis was performed. The purpose of stepwise regression is to identify the subset of predictors that best explain the response variable. In this case, the response variables are TARGET_AMT, KIDSDRIV, AGE, HOMEKIDS, YOJ, INCOME, HOME_VAL, TRAVTIME, BLUEBOOK, TIF, OLDCLAIM, CLM_FREQ, MVR_PTS, and CAR_AGE. The model selects the best subset of these predictors by iteratively adding or removing variables based on statistical criteria, in this case the Akaike Information Criterion (AIC).

The final model contains the variables TARGET_AMT, KIDSDRIV, AGE, HOMEKIDS, YOJ, HOME_VAL, TRAVTIME, BLUEBOOK, TIF, OLDCLAIM, CLM_FREQ ,MVR_PTS, CAR_AGE. This subset of variables yielded the lowest AIC value among all possible subsets of variables considered by the algorithm.

To evaluate the performance of a machine learning model, it is important to split the data into training and test sets. The training set is used to train the model, while the test set is used to evaluate how well the model generalizes to new, unseen data.

```{r seung-model, warning=FALSE }
library(MASS)

set.seed(333)
trainIndex <- sample(1:nrow(trainClean), round(0.7*nrow(trainClean)), replace = FALSE)
trainData <- trainClean[trainIndex, ]
testData <- trainClean[-trainIndex, ]


trainData <- na.omit(trainData)
testData <- na.omit(testData)


logit <- glm(formula = TARGET_FLAG ~ . - TARGET_AMT, data=rawData, family = "binomial" (link="logit"))

summary(logit)
```




The model's coefficient of 0.5860083 represents the correlation between the predicted values and actual values of the target variable, which is the probability of a car insurance claim being filed (TARGET_FLAG). A value of 0.60 indicates a moderate positive correlation between the predicted and actual values.

The model's coefficients for INCOME and HOME_VAL are negative, indicating that as these variables decrease, the predicted probability of a car insurance claim being filed (TARGET_FLAG) increases. This observation seems to contradict the expectation that individuals with higher incomes and more valuable homes would be safer drivers and thus less likely to file insurance claims.

If the result is counter intuitive, it might be worthwhile to conduct another experiment to verify the accuracy of the outcome.

```{r}
exp(logit$coefficients)
```

```{r}
logitscalar <- mean(dlogis(predict(logit, type = "link")))
logitscalar * coef(logit)
```


```{r}
confint.default(logit)
```


```{r}
predlogit <- predict(logit, type="response")
rawData2$pred1 <- predict(logit, type="response")
summary(predlogit)
```
```{r}
table(true = rawData$TARGET_FLAG, pred = round(fitted(logit)))
```

```{r}
#plots for Seung's Model logit
par(mfrow=c(2,2))
plot(logit)
```


```{r}
data.frame(rawData2$pred1) %>%
    ggplot(aes(x = rawData2.pred1)) + 
    geom_histogram(bins = 50, fill = 'green') +
    labs(title = 'Histogram of Predictions') +
    theme_bw()
```
```{r}
plot.roc(rawData$TARGET_FLAG, rawData2$pred1)

#extract variables that are significant and rerun model
sigvars <- data.frame(summary(logit)$coef[summary(logit)$coef[,4] <= .05, 4])
sigvars <- add_rownames(sigvars, "vars")
```



## Anthony's First Model logit2

Of the models we build, we will build our first two models after an exhaustive outlier cleaning. It seems that both models favor managers for a lower insurance rate, interestingly enough. Additionally, the type of car has a significant impact on the amount. Also, both models agree that married couples are charged significantly lower rates, whereas this is even worse in the event that an unmarried person is also a single parent. Finally, both models agree that it is extremely significant whether the drivers are in urban or rural areas, such that urban drivers get charged around $1500 more than their rural counterparts.

```{r a-model1, eval=FALSE, warning=FALSE}
colist<-dplyr::pull(sigvars, vars)
# colist<-colist[2:11]
colist<-c("KIDSDRIV","INCOME","PARENT1","HOME_VAL","MSTATUS","EDUCATION","JOB","TRAVTIME","CAR_USE","BLUEBOOK","TIF","CAR_TYPE","CLM_FREQ","REVOKED","MVR_PTS","URBANICITY")

idx <- match(colist, names(rawData))
rawDatamod2 <- cbind(rawData[,idx], rawData2['TARGET_FLAG'])


logit2 <- glm(TARGET_FLAG ~ ., data=rawDatamod2, family = "binomial" (link="logit"))
summary(logit2)
```

```{r a-model2, eval=FALSE, warning=FALSE}
exp(logit2$coefficients)
```

```{r a-model3, eval=FALSE, warning=FALSE}
logit2scalar <- mean(dlogis(predict(logit2, type = "link")))
logit2scalar * coef(logit2)
```

```{r }
predlogit2 <- predict(logit2, type="response")
rawData2$pred2 <- predict(logit2, type="response")
```


## Additional Transforms

Since we created our initial models, we decided that there are three additional transforms we want to add. We believe that having a claim is predictive of making a claim again, so we will have a boolean column for any number of claims. Additionally, we decided that the only two relevant job types are doctors and managers, based on the models. Therefore, we will also make two boolean columns for if they are a manager or doctor.

```{r tf-ins}
### Creates a factor column for Manager/Doctor jobs and no claim history, specific to insurance dataset
InsuranceTransforms <- function(df) {
  df %>%
    dplyr::mutate(
      IS_MANAGER = as.factor(as.integer(JOB %in% "Manager")),
      IS_DOCTOR = as.factor(as.integer(JOB %in% "Doctor")),
      NO_CLM_HIST = as.factor(as.integer(CLM_FREQ < 1))
    ) %>%
    dplyr::select(-JOB, -CLM_FREQ)
}
```

## Further Notes

Although new models were attempted with the new transforms and additional considerations, none of them successfully beat the existing model by a statistically significant margin. We compared each of the models using ANOVA and evaluated the LM models with MAE and RMSE, whereas the logistical models were compared using F1 and Accuracy metrics. Overall, we kept the previously accepted models due to the lack of advancement.

```{r }
summary(predlogit2)
```

```{r more-tests, eval=FALSE}
table(true = rawData$TARGET_FLAG, pred = round(fitted(logit2)))

```

Anthony's First Model plots for logit2

```{r test-excess, eval=FALSE}
par(mfrow=c(2,2))
plot(logit2)
```

```{r lm-assist, eval=FALSE}
data.frame(rawData2$pred2) %>%
    ggplot(aes(x = rawData2.pred2)) + 
    geom_histogram(bins = 50, fill = 'blue') +
    labs(title = 'Histogram of Predictions') +
    theme_bw()
```
```{r}
plot.roc(rawData$TARGET_FLAG, rawData2$pred2)
```



# IvanTikhonov Best Model

```{r}
#PC Model no racial bias
logit3 <- glm(TARGET_FLAG ~ KIDSDRIV + INCOME + HOME_VAL + TRAVTIME, data=rawData, family = "binomial" (link="logit"))
summary(logit3)
```
```{r}
exp(logit3$coefficients)
```

```{r}
predlogit3 <- predict(logit3, type="response")
rawData2$pred3 <- predict(logit3, type="response")
summary(predlogit3)
```
```{r}
table(true = rawData$TARGET_FLAG, pred = round(fitted(logit3)))
```
```{r}
par(mfrow=c(2,2))
plot(logit3)
```
```{r}
data.frame(rawData2$pred3) %>%
    ggplot(aes(x = rawData2.pred3)) + 
    geom_histogram(bins = 50, fill = 'red') +
    labs(title = 'Histogram of Predictions') +
    theme_bw()
```

```{r}
plot.roc(rawData$TARGET_FLAG, rawData2$pred3)
```

```{r}
logit3scalar <- mean(dlogis(predict(logit3, type = "link")))
logit3scalar * coef(logit3)
```
```{r}
round(logitscalar * coef(logit),2)
```

```{r}
round(logit2scalar * coef(logit2),2)
```

```{r}
round(logit3scalar * coef(logit3),2)
```

## Seung's Model

```{r}
model <- lm(TARGET_AMT ~ ., data=rawData)
summary(model)
```

```{r}
par(mfrow=c(1,2))
plot(model$residuals ~ model$fitted.values)
plot(model$fitted.values,rawData$TARGET_AMT)
```

```{r}
par(mfrow=c(2,2))
plot(model)
```

```{r warning=FALSE}
#extract variables that are significant and rerun model
sigvars <- data.frame(summary(model)$coef[summary(model)$coef[,4] <= .05, 4])
sigvars <- add_rownames(sigvars, "vars")
```

```{r}
colist<-dplyr::pull(sigvars, vars)
colist<-c("TARGET_FLAG","BLUEBOOK","REVOKED","MVR_PTS","CAR_AGE")

idx <- match(colist, names(rawData))
rawDatamod2 <- cbind(rawData[,idx], rawData['TARGET_AMT'])
```


## Anthony's First Model

```{r}
model2<-lm(TARGET_AMT ~ ., data=rawDatamod2)

summary(model2)
```
```{r}
par(mfrow=c(2,2))
plot(model2$residuals ~ model2$fitted.values)
plot(model2$fitted.values,rawData$TARGET_AMT)


par(mfrow=c(2,2))
```
```{r}
plot(model2)
```

```{r}
par(mfrow=c(1,2))
plot(model2$residuals ~ model2$fitted.values, main="New Reduced Var Model")
abline(h = 0)
plot(model$residuals ~ model$fitted.values, main="Orignal Model All Vars")
abline(h = 0)
```
# IvanTikhonov Model
```{r}
#remove variables with opposite coefficients

model3<-lm(TARGET_AMT ~ KIDSDRIV + INCOME + HOME_VAL + TRAVTIME, data=rawData)
summary(model3)
```

```{r}
par(mfrow=c(1,2))
plot(model3$residuals ~ model3$fitted.values)
plot(model3$fitted.values,rawData$TARGET_AMT)
```
```{r}
par(mfrow=c(2,2))
plot(model3)
```

##  Charles Ugiagbe  Select Models

```{r}
rawTestData2<- rawTestData
```

```{r warning=FALSE}
rawTestData = read.csv(file="https://raw.githubusercontent.com/IvanGrozny88/DATA-621-HW-4/main/insurance-evaluation-data.csv")

rawTestData$TARGET_AMT <- 0
rawTestData$TARGET_FLAG <- 0


rawTestData = as.tbl(rawTestData) %>% 
  mutate_at(c("INCOME","HOME_VAL","BLUEBOOK","OLDCLAIM"),
            currencyconv) %>% 
  mutate_at(c("EDUCATION","JOB","CAR_TYPE","URBANICITY"),
            underscore) %>% 
  mutate_at(c("EDUCATION","JOB","CAR_TYPE","URBANICITY"),
            as.factor) %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG))

# impute data for missing values
# use column mean for calculation

rawTestData$HOMEKIDS <- log(rawTestData$HOMEKIDS+1)
rawTestData$MVR_PTS <- log(rawTestData$MVR_PTS+1)
rawTestData$OLDCLAIM <- log(rawTestData$OLDCLAIM+1)
rawTestData$TIF <- log(rawTestData$TIF+1)
rawTestData$KIDSDRIV <- log(rawTestData$KIDSDRIV+1)
rawTestData$CLM_FREQ <- log(rawTestData$CLM_FREQ+1)

# use column mean for calculation

rawTestData$AGE[is.na(rawTestData$AGE)] <- mean(rawTestData$AGE, na.rm=TRUE)
rawTestData$YOJ[is.na(rawTestData$YOJ)] <- mean(rawTestData$YOJ, na.rm=TRUE)
rawTestData$HOME_VAL[is.na(rawTestData$HOME_VAL)] <- mean(rawTestData$HOME_VAL, na.rm=TRUE)
rawTestData$CAR_AGE[is.na(rawTestData$CAR_AGE)] <- mean(rawTestData$CAR_AGE, na.rm=TRUE)

rawTestData$INCOME[is.na(rawTestData$INCOME)] <- mean(rawTestData$INCOME, na.rm=TRUE)

#get complete cases


#remove rad per correlation in prior section

rawTestData <- rawTestData[, !(colnames(rawTestData) %in% c("INDEX"))]



TARGET_FLAG <- predict(logit, newdata = rawTestData, type="response")

y_pred_num <- ifelse(TARGET_FLAG > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
summary(y_pred)
```

```{r}
rbind(round(summary(predlogit),4), round(summary(TARGET_FLAG),4)) %>% kable()
```

```{r}
rawTestData$TARGET_FLAG <- as.factor(rawTestData$TARGET_FLAG)

rawTestData2 <- rawTestData[, !(colnames(rawTestData) %in% c("TARGET_FLAG"))]
TARGET_AMT<- predict(model, newdata = rawTestData, interval='confidence') #data from scaling originally to get to actual wins
summary(TARGET_AMT)
```
```{r}
summary(model)
```

```{r}
logit4 <- glm(TARGET_FLAG ~ KIDSDRIV + INCOME + HOME_VAL + TRAVTIME, data=rawData, family = "binomial" (link="logit"))
summary(logit4)
```
```{r}
exp(logit4$coefficients)
```

```{r}
predlogit4 <- predict(logit4, type="response")
rawData2$pred4 <- predict(logit4, type="response")
summary(predlogit4)
```

```{r}
table(true = rawData$TARGET_FLAG, pred = round(fitted(logit4)))
```

```{r}
par(mfrow=c(2,2))
plot(logit4)
```

```{r}
data.frame(rawData2$pred3) %>%
    ggplot(aes(x = rawData2.pred4)) + 
    geom_histogram(bins = 50, fill = 'violet') +
    labs(title = 'Histogram of Predictions') +
    theme_bw()
```

```{r}
plot.roc(rawData$TARGET_FLAG, rawData2$pred4)
```





# Conclusions and Suggestions

Overall, none of our models really seemed to have the desired correlation that we were looking for. The one model that Seung made predicts well on the condition that the regression model can correctly predict claim amount. But since we don't have such a pristine model available, both models perform poorly. Perhaps we need to look into more data engineering techniques to better normalize our data. We spent an excessive amount of time doing imputations and statistical testing but it was overall ineffective to producing results.

# References

Data is provided by CUNY School of Professional Studies and is thus sourced by open data initiatives.

# Appendix

```{r, echo=TRUE, class.output="foldable"}
CleanInsuranceStrings
```

```{r, echo=TRUE, class.output="foldable"}
InsuranceTransforms
```


